# When to stop

```{r stop load and select data}
# load data
stop <- read_csv("../../data/processed/stop.csv")
stop_breaks <- read_csv("../../data/processed/stop_breaks.csv")
```

## Main analysis

In this section, we analyze when players decided to stop. For this analysis, players need to have 5 wins and 5 losses (all sessions combined) to be included in the analysis. The table below shows the number of players and rounds left in this analysis.

```{r stop - data preparation, results='asis'}
# players need to have at least 5 wins and 5 losses
players_incldue <- stop %>%
  filter(total_count >= 5) %>%
  pivot_wider(
    id_cols = c(group, id_player), names_from = outcome,
    values_from = stop_prob
  ) %>%
  drop_na() %>%
  .$id_player

stop <- stop %>% filter(id_player %in% players_incldue)
stop_breaks <- stop_breaks %>% filter(id_player %in% players_incldue)
  
# calculate the remaining number of players and rounds
count <- stop %>%
  group_by(group, id_player) %>%
  summarize(nrounds = sum(total_count)) %>%
  group_by(group) %>%
  summarize(
    player = n(),
    round = mean(nrounds),
    sd = sd(nrounds),
    min = min(nrounds),
    max = max(nrounds)
  )

kable(count, caption = "Number of players and rounds left") %>%
  kable_styling()
```

### Plot

For the remaining players, we first plot the probabilities of ending a session after a win versus after a loss.

(ref:stop-prob-plot-cap) Probability of ending a session after a win and a loss. Error bars stand for 95\% within-subject confidence intervals.

```{r stop plot, fig.cap="(ref:stop-prob-plot-cap)", fig.width=3.5, fig.height=3.5}
stop_summary <- stop %>%
  group_by(group) %>%
  group_modify(~Rmisc::summarySEwithin(.x, 
                                       measurevar = "stop_prob", 
                                       idvar = "id_player",
                                       withinvars = "outcome")) %>%
  mutate(
    outcome = factor(outcome, levels = c("loss", "win")),
    group = factor(group, levels = c("High-Risk", "Low-Risk"),
                   labels = c("High-Involvement", "Low-Involvement"))
  )

# plot the probabilities of stopping a session after a win vs. a loss
stop_plot <- ggplot(stop_summary, 
                    aes(outcome, stop_prob, color = group, shape = group)) +
  geom_point(position = position_dodge(width = dodge_width), size = shape_size) +
  geom_line(aes(group = group, linetype = group), 
            position = position_dodge(width = dodge_width)) +
  geom_errorbar(aes(ymin = stop_prob-ci, ymax = stop_prob+ci), width = 0.2,
                position = position_dodge(width = dodge_width)) +
  # use colorblind-friendly color
  scale_color_manual(values=color_values) + 
  scale_x_discrete(labels=c("loss" = "Loss", "win" = "Win")) +
  labs(x = "Previous outcome", 
       y = "Probability of ending a session (%)",
       color = "Group", shape = "Group", linetype = "Group") +
  theme(legend.position = "top", legend.direction = "horizontal") 

stop_plot

ggsave("figure/stop.pdf", stop_plot, width = 3.5, height = 3.5)
```

The high-involvement players overall had a lower probability of ending a session, as they tended to have longer sessions on average. To control for this overall between-group difference in the stopping probability, we divided the probability of stopping after a win or after a loss by the overall probability of stopping for each player.

(ref:stop-prob-controlled-plot-cap) Probability of ending a session after a win and a loss, controlling for the overall probability of stopping. Error bars stand for 95\% within-subject confidence intervals.

```{r stop controlled plot, fig.cap="(ref:stop-prob-controlled-plot-cap)", fig.width=4, fig.height=4}
stop <- stop %>%
  mutate(stop_prob_relative = stop_prob/stop_overall)

stop_summary <- stop %>%
  group_by(group) %>%
  group_modify(~Rmisc::summarySEwithin(.x, measurevar = "stop_prob_relative", 
                                       idvar = "id_player", withinvars = "outcome")) %>%
  mutate(outcome = factor(outcome, levels = c("loss", "win")),
         group = factor(group, levels = c("High-Risk", "Low-Risk"),
                        labels = c("High-Involvement", "Low-Involvement")))

# plot winning probabilities
stop_plot <- ggplot(stop_summary, 
                    aes(outcome, stop_prob_relative, color = group, shape = group)) +
  geom_point(position = position_dodge(width = dodge_width), size = shape_size) +
  geom_line(aes(group = group, linetype = group), 
            position = position_dodge(width = dodge_width)) +
  geom_errorbar(aes(ymin = stop_prob_relative-ci, ymax = stop_prob_relative+ci), 
                width = 0.2,
                position = position_dodge(width = dodge_width)) +
  # use colorblind-friendly color
  scale_color_manual(values=color_values) + 
  scale_x_discrete(labels=c("loss" = "Loss", "win" = "Win")) +
  labs(x = "Previous outcome", 
       y = "Relative likelihood of ending a session",
       color = "Group", shape = "Group", linetype = "Group") +
  theme(legend.position = "top", legend.direction = "horizontal") 

stop_plot

# this will be the main figure included in the paper
stop_plot_main <- stop_plot

ggsave("figure/stop_relative.pdf", stop_plot, width = 4, height = 4)
```

### ANOVA

The stopping probabilities are analyzed with a mixed ANOVA, with the previous outcome (loss vs. win, within subjects) and the involvement level (high vs. low, between subjects) as independent variables. we conduct the analysis for both the raw probabilities of stopping, and the relative likelihood of stopping after taking the overall probability of stopping into account.

```{r stop anova analysis, cache = TRUE, results = "asis"}
# run an ANOVA
stop_anova <- aov_ez(id = "id_player"
                     , dv = "stop_prob"
                     , within = "outcome"
                     , between = "group"
                     , data = stop)

stop_anova <- nice(stop_anova)

# do a bit formatting
stop_anova <- stop_anova %>%
  mutate(
    Effect = c("Involvement Level", "Outcome", "Interaction"),
    # remove asterisks from the F values
    `F` = str_remove_all(`F`, "[*]"),
    # remove white spaces
    `F` = str_trim(`F`)
  ) %>%
  rename(p = "p.value")

kable(stop_anova, 
      caption = "When do players decide to stop?",
      booktabs = T, align = "l")
```


```{r relative stop anova analysis, cache = TRUE, results = "asis"}
# run an ANOVA
stop_anova <- aov_ez(id = "id_player"
                     , dv = "stop_prob_relative"
                     , within = "outcome"
                     , between = "group"
                     , data = stop)

stop_anova <- nice(stop_anova)

# do a bit formatting
stop_anova <- stop_anova %>%
  mutate(
    Effect = c("Involvement Level", "Outcome", "Interaction"),
    # remove asterisks from the F values
    `F` = str_remove_all(`F`, "[*]"),
    # remove whitespace
    `F` = str_trim(`F`)
  ) %>%
  rename(p = "p.value")

kable(stop_anova, 
      caption = "When do players decide to stop? (Controlling for the overall probability of stopping)",
      booktabs = T, align = "l")
```


### Pairwise comparisons

Next we conduct a series of pairwise comparisons. First, we examine the effects within groups, by comparing the probability of stopping after a loss vs. that after a win, for each group separately. Next we examine the effects between groups, by calculating a difference score for each player (stopping after a loss - stopping after a win) and compare the difference scores between two groups. P values are corrected for multiple comparisons using the Holm-Bonferroni method.

```{r stop pairwise t tests, results='asis'}
# for pairwise comparisons, we need the data in the wide format
stop_wide <- stop %>%
  pivot_wider(id_cols = c(group, id_player),
              names_from = outcome,
              values_from = stop_prob) %>%
  # calculate a difference score
  mutate(diff = loss - win)

# within-group comparisons
# first, compare the effect within each group
comp_within <- stop_wide %>%
  group_by(group) %>%
  group_modify(~TES(.x$loss, .x$win, paired = TRUE)) %>%
  mutate(d = dz, g = gav) %>%
  ungroup() %>% select(-group, -dz, -gav, -dav)

# between group comparisons
comp_between1 <- TES(filter(stop_wide, group == "High-Risk")$loss,
                    filter(stop_wide, group == "Low-Risk")$loss, paired = FALSE)

comp_between2 <- TES(filter(stop_wide, group == "High-Risk")$win,
                    filter(stop_wide, group == "Low-Risk")$win, paired = FALSE)

comp_between3 <- TES(filter(stop_wide, group == "High-Risk")$diff,
                    filter(stop_wide, group == "Low-Risk")$diff, paired = FALSE)

comp_between <- bind_rows(comp_between1, comp_between2, comp_between3)

# combine and correct p values for multiple comparisons
comp <- bind_rows(comp_within, comp_between) %>%
  mutate(
    Comparison = c("High-Loss vs. High-Win", 
                   "Low-Loss vs. Low-Win", 
                   "High-Loss vs. Low-Loss",
                   "High-Win vs. Low-Win",
                   "(High-Loss - High-Win) vs. (Low-Loss - Low-Win)"),
    pt = p.adjust(pt, method = "holm")
  ) 

# do some formatting
comp <- comp %>%
  mutate(
    across(mean_x:upperCI, ~round(.x, digits = 2)),
    across(c(d, g), ~round(.x, digits = 3)),
    t = round(t, digits = 1),
    df = round(df, digits = 1),
    p = ifelse(pt < .001, "<.001", sprintf("%.3f", round(pt, digits = 3))),
    logBF = round(logBF, digits = 2)
  ) %>%
  select(Comparison, mean_x:t, p, logBF, g) 

# show the table
kable(comp, caption = "Pairwise comparisons on when to stop",
      booktabs = T, align = "l") %>%
  scroll_box(width = "100%", height = "100%")

```

Repeat the pairwise comparisons, but this time accounting for the overall probability of stopping.

```{r relative stop pairwise t tests, results='asis'}
# for pairwise comparisons, we need the data in the wide format
stop_wide <- stop %>%
  pivot_wider(id_cols = c(group, id_player),
              names_from = outcome,
              values_from = stop_prob_relative) %>%
  # calculate a difference score
  mutate(diff = loss - win)

# within-group comparisons
# first, compare the effect within each group
comp_within <- stop_wide %>%
  group_by(group) %>%
  group_modify(~TES(.x$loss, .x$win, paired = TRUE)) %>%
  mutate(d = dz, g = gav) %>%
  ungroup() %>% select(-group, -dz, -gav, -dav)

# between group comparisons
comp_between1 <- TES(filter(stop_wide, group == "High-Risk")$loss,
                    filter(stop_wide, group == "Low-Risk")$loss, paired = FALSE)

comp_between2 <- TES(filter(stop_wide, group == "High-Risk")$win,
                    filter(stop_wide, group == "Low-Risk")$win, paired = FALSE)

comp_between3 <- TES(filter(stop_wide, group == "High-Risk")$diff,
                    filter(stop_wide, group == "Low-Risk")$diff, paired = FALSE)

comp_between <- bind_rows(comp_between1, comp_between2, comp_between3)

# combine and correct p values for multiple comparisons
comp <- bind_rows(comp_within, comp_between) %>%
  mutate(
    Comparison = c("High-Loss vs. High-Win", 
                   "Low-Loss vs. Low-Win", 
                   "High-Loss vs. Low-Loss",
                   "High-Win vs. Low-Win",
                   "(High-Loss - High-Win) vs. (Low-Loss - Low-Win)"),
    pt = p.adjust(pt, method = "holm")
  ) 

# do some formatting
comp <- comp %>%
  mutate(
    across(mean_x:upperCI, ~round(.x, digits = 2)),
    across(c(d, g), ~round(.x, digits = 3)),
    t = round(t, digits = 1),
    df = round(df, digits = 1),
    p = ifelse(pt < .001, "<.001", sprintf("%.3f", round(pt, digits = 3))),
    logBF = round(logBF, digits = 2)
  ) %>%
  select(Comparison, mean_x:t, p, logBF, g) 

# show the table
kable(comp, caption = "Pairwise comparisons on when to stop",
      booktabs = T, align = "l") %>%
  scroll_box(width = "100%", height = "100%")


```

## With breaks

Here we repeat the analysis above (controlling for the overall probability of stopping), while also taking breaks in play into account.

### Plot

(ref:stop-prob-controlled-breaks-plot-cap) Probability of ending a session after a win and a loss, controlling for the overall probability of stopping. Error bars stand for 95\% within-subject confidence intervals.

```{r stop controlled plot with breaks, fig.cap="(ref:stop-prob-controlled-breaks-plot-cap)", fig.width=4, fig.height=4}
stop_breaks <- stop_breaks %>%
  mutate(stop_prob_relative = stop_prob/stop_overall)

stop_summary <- stop_breaks %>%
  group_by(group) %>%
  group_modify(~Rmisc::summarySEwithin(.x, measurevar = "stop_prob_relative", 
                                       idvar = "id_player", withinvars = "outcome")) %>%
  mutate(
    outcome = factor(outcome, levels = c("loss", "win")),
    group = factor(group, levels = c("High-Risk", "Low-Risk"),
                   labels = c("High-Involvement", "Low-Involvement"))
  )

# plot winning probabilities
stop_plot <- ggplot(stop_summary, 
                    aes(outcome, stop_prob_relative, color = group, shape = group)) +
  geom_point(position = position_dodge(width = dodge_width), size = shape_size) +
  geom_line(aes(group = group, linetype = group), 
            position = position_dodge(width = dodge_width)) +
  geom_errorbar(aes(ymin = stop_prob_relative-ci, ymax = stop_prob_relative+ci), 
                width = 0.2,
                position = position_dodge(width = dodge_width)) +
  # use colorblind-friendly color
  scale_color_manual(values=color_values) + 
  scale_x_discrete(labels=c("loss" = "Loss", "win" = "Win")) +
  labs(x = "Previous outcome", 
       y = "Relative likelihood of stopping",
       color = "Group", shape = "Group", linetype = "Group") +
  theme(legend.position = "top", legend.direction = "horizontal") 

stop_plot

ggsave("figure/stop_breaks.pdf", stop_plot, width = 4, height = 4)
```

### ANOVA

```{r relative stop anova analysis with breaks, cache = TRUE, results = "asis"}
# run an ANOVA
stop_anova <- aov_ez(id = "id_player"
                     , dv = "stop_prob_relative"
                     , within = "outcome"
                     , between = "group"
                     , data = stop_breaks)

stop_anova <- nice(stop_anova)

# do a bit formatting
stop_anova <- stop_anova %>%
  mutate(
    Effect = c("Involvement Level", "Outcome", "Interaction"),
    # remove asterisks from the F values
    `F` = str_remove_all(`F`, "[*]"),
    # remove whitespace
    `F` = str_trim(`F`)
  ) %>%
  rename(p = "p.value")

kable(stop_anova, 
      caption = "When do players decide to stop?",
      booktabs = T, align = "l")
```

### Pairwise comparisons

```{r relative stop with breaks pairwise t tests, results='asis'}
# for pairwise comparisons, we need the data in the wide format
stop_wide <- stop_breaks %>%
  pivot_wider(id_cols = c(group, id_player),
              names_from = outcome,
              values_from = stop_prob_relative) %>%
  # calculate a difference score
  mutate(diff = loss - win)

# within-group comparisons
# first, compare the effect within each group
comp_within <- stop_wide %>%
  group_by(group) %>%
  group_modify(~TES(.x$loss, .x$win, paired = TRUE)) %>%
  mutate(d = dz, g = gav) %>%
  ungroup() %>% select(-group, -dz, -gav, -dav)

# between group comparisons
comp_between1 <- TES(filter(stop_wide, group == "High-Risk")$loss,
                    filter(stop_wide, group == "Low-Risk")$loss, paired = FALSE)

comp_between2 <- TES(filter(stop_wide, group == "High-Risk")$win,
                    filter(stop_wide, group == "Low-Risk")$win, paired = FALSE)

comp_between3 <- TES(filter(stop_wide, group == "High-Risk")$diff,
                    filter(stop_wide, group == "Low-Risk")$diff, paired = FALSE)

comp_between <- bind_rows(comp_between1, comp_between2, comp_between3)

# combine and correct p values for multiple comparisons
comp <- bind_rows(comp_within, comp_between) %>%
  mutate(
    Comparison = c("High-Loss vs. High-Win", 
                   "Low-Loss vs. Low-Win", 
                   "High-Loss vs. Low-Loss",
                   "High-Win vs. Low-Win",
                   "(High-Loss - High-Win) vs. (Low-Loss - Low-Win)"),
    pt = p.adjust(pt, method = "holm")
  ) 

# do some formatting
comp <- comp %>%
  mutate(
    across(mean_x:upperCI, ~round(.x, digits = 2)),
    across(c(d, g), ~round(.x, digits = 3)),
    t = round(t, digits = 1),
    df = round(df, digits = 1),
    p = ifelse(pt < .001, "<.001", sprintf("%.3f", round(pt, digits = 3))),
    logBF = round(logBF, digits = 2)
  ) %>%
  select(Comparison, mean_x:t, p, logBF, g) 

# show the table
kable(comp, caption = "Pairwise comparisons on when to stop",
      booktabs = T, align = "l") %>%
  scroll_box(width = "100%", height = "100%")

```
