# Process data

The raw data on the individual game level cannot be shared as they are proprietary. We nevertheless provide the code that we used to process the raw data to get processed player-level data. The processed (summary) data are publicly shared (in the `data/processed` folder) and can be used to reproduce the results that we report in the paper.


```{r load-raw-data}
# load data
load("../../data/processed/d_low_risk.RData")
load("../../data/processed/d_high_risk.RData")

# clean and combine data
d_low_risk <- d_low_risk %>%
  mutate(
    # add group name,
    group = "Low-Risk",
    # low-risk players each has a risk level of 0
    level = 0,
    # add "LR" (low-risk) to the player id
    id_player = str_c("LR", id_player, sep = "_")
  ) %>%
  select(group, level, everything())

d_high_risk <- d_high_risk %>%
  mutate(
    # add group name,
    group = "High-Risk",
    # add "HR" (high-risk) to the player id
    id_player = str_c("HR", id_player, sep = "_")
  ) %>%
  select(group, level, everything())

# combine data from both groups
d <- rbind(d_high_risk, d_low_risk)

# remove the original two data frames
remove(d_high_risk)
remove(d_low_risk)
```

## Player and play behavior

### descriptive_data.csv

First, we process the data to get descriptive data on the players and their general play behaviors.

The `descriptive_data.csv` file contains the following variables.

* group: the group to which a player belongs, High-Risk or Low-Risk.
* id_player: the anonymous id of a player.
* age: age of a player.
* gender: gender of a player, Female or Male.
* income: annual income of a player, in euro. Players without income information are denoted NA.
* risk_0: whether the player receives a risk level of 0 in the data, 1 = yes, 0 = no.
* risk_3: whether the player receives a risk level of 3 in the data, 1 = yes, 0 = no.
* risk_4: whether the player receives a risk level of 4 in the data, 1 = yes, 0 = no.
* risk_5: whether the player receives a risk level of 5 in the data, 1 = yes, 0 = no.
* n_session: the total number of sessions played.
* n_round_total: the total number of rounds played.
* n_round_mean: the mean number of rounds played per session.
* n_round_median: the median number of rounds played per session.
* bet_max: the maximum bet size across all rounds, in euro.
* bet_min: the minimum bet size across all rounds, in euro.
* bet_mean: the mean bet size across all rounds, in euro.
* bet_median: the median bet size across all rounds, in euro.
* win_prob: the probability of winning (%) across all rounds.
* win_max: the maximum amount of win on rounds when players win, in euro. Win defined as the win amount minus the stake amount.
* win_min: the minimum amount of win on rounds when players win, in euro.
* win_mean: the mean amount of win on rounds when players win, in euro.
* win_median:  the median amount of win on rounds when players win, in euro.
* loss_max: the maximum amount of loss on rounds when players lose, in euro. Loss amount equals the stake amount.
* loss_min: the minimum amount of loss on rounds when players lose, in euro.
* loss_mean: the mean amount of loss on rounds when players lose, in euro.
* loss_median: the median amount of loss on rounds when players lose, in euro.
* total_spent: the total amount of money spent (i.e., lost), in euro. A negative value indicates that players overall won money.

```{r descriptive_data}
# get the age, gender and income of each player
age_gender <- d %>%
  group_by(id_player) %>%
  sample_n(1) %>%
  mutate(gender = recode(gender, F = "Female", M = "Male")) %>%
  select(group, id_player, age, gender, income)

# get the risk level(s) of each player
risk_level_0 <- d %>%
  filter(level == 0) %>%
  .$id_player %>% unique

risk_level_3 <- d %>%
  filter(level == 3) %>%
  .$id_player %>% unique

risk_level_4 <- d %>%
  filter(level == 4) %>%
  .$id_player %>% unique

risk_level_5 <- d %>%
  filter(level == 5) %>%
  .$id_player %>% unique

risk_levels <- age_gender %>%
  mutate(
    risk_0 = ifelse(id_player %in% risk_level_0, 1, 0),
    risk_3 = ifelse(id_player %in% risk_level_3, 1, 0),
    risk_4 = ifelse(id_player %in% risk_level_4, 1, 0),
    risk_5 = ifelse(id_player %in% risk_level_5, 1, 0)
  ) %>%
  select(id_player, risk_0, risk_3, risk_4, risk_5)

# count the number of sessions and rounds for each player
rounds <- d %>%
  # count the number of rounds in each session
  count(id_player, id_session) %>%
  # count for each player:
  group_by(id_player) %>%
  summarize(
    n_session = n(), # the total number of sessions
    n_round_total = sum(n), # the total number of rounds (all sessions combined)
    n_round_mean =  mean(n), # the mean number of rounds per session
    n_round_median = median(n) # the median number of rounds per session
  ) %>%
  select(id_player, n_session, n_round_total, n_round_mean, n_round_median)

# for each player, calculate the max, min, mean and median of stakes across all rounds
bets <- d %>%
  group_by(id_player) %>%
  summarize(
    bet_max = max(stake),
    bet_min = min(stake),
    bet_mean = mean(stake),
    bet_median = median(stake)
  ) %>%
  select(id_player, bet_max, bet_min, bet_mean, bet_median)

# for each player, calculate their winning probability, the min, max, mean and median win amount
win_prob <- d %>%
  group_by(id_player) %>%
  summarize(win_prob = mean(win > 0) * 100) %>%
  select(id_player, win_prob)

wins <- d %>%
  # select winning rounds
  filter(win > 0) %>%
  # for each player, calculate the max, min, mean and median win amount
  # the real win amount is the win amount displayed by the game, minus
  # the stake
  mutate(win = win - stake) %>%
  group_by(id_player) %>%
  summarize(
    win_max = max(win),
    win_min = min(win),
    win_mean = mean(win),
    win_median = median(win)
  ) %>%
  select(id_player, win_max, win_min, win_mean, win_median)

# on the rounds where players lost
# calculate their max, min, mean, median and total loss amount
losses <- d %>%
  # select losing rounds
  filter(win == 0) %>%
  # loss the same as the stake
  mutate(loss = stake) %>%
  group_by(id_player) %>%
  summarize(
    loss_max = max(loss),
    loss_min = min(loss),
    loss_mean = mean(loss),
    loss_median = median(loss),
  ) %>%
  select(id_player, loss_max, loss_min, loss_mean, loss_median)

# for each player, calculate the total spent
total_spent <- d %>%
  mutate(spent = stake - win) %>%
  group_by(id_player) %>%
  summarize(total_spent = sum(spent))

# put all data frames together
d_descriptive <- age_gender %>%
  full_join(risk_levels, by = "id_player") %>%
  full_join(rounds, by = "id_player") %>%
  full_join(bets, by = "id_player") %>%
  full_join(win_prob, by = "id_player") %>%
  full_join(wins, by = "id_player") %>%
  full_join(losses, by = "id_player") %>%
  full_join(total_spent, by = "id_player")

# save the data frames as a csv file
write_csv(d_descriptive, "../../data/processed/descriptive_data.csv")

# remove data frames
remove(age_gender, risk_levels, rounds, bets, win_prob, wins, losses, total_spent, d_descriptive)
```

### sessions_count.csv

To further characterize the play behavior, we count the number of rounds played in each session. For the high-risk and low-risk groups separately, we count the number of sessions that contained certain number of rounds (combined across all players in the same group). The data are saved in the `sessions_count.csv` file, which contains the following variables.

* group: the group to which players belong, High-Risk or Low-Risk.
* n_round_bin: the number of rounds per session is divided into bins. (a, b] means a certain session has > a and <= b rounds. Note that bin widths become larger for larger values.
* n_round_label: a numeric label for each bin, from 1 till 34 (included for later plotting).
* n_session: the number of sessions that contain certain numbers of rounds, combined across all players in the same group. E.g, for the High-Risk group, 13314 sessions fall into the (0,10] bin, meaning that for all players in the High-Risk group, 13314 sessions contain > 0 and <= 10 rounds.


```{r sessions_count}
# we count the number of sessions that contain certain number of rounds
# across all players, for the high-risk and low-risk groups separately
sessions_count <- d %>%
  # count the number of rounds in each session
  count(group, id_player, id_session) %>%
  # divide the number of rounds into into bins
  mutate(
    n_round_bin = cut(n, breaks = c((0:20)*10, (3:10)*100, (2:7)*1000)),
    n_round_label = cut(n, breaks = c((0:20)*10, (3:10)*100, (2:7)*1000), labels = 1:34)
  ) %>%
  # calculate the number of sessions with certain numbers of rounds
  count(group, n_round_bin, n_round_label, name = "n_session") %>%
  mutate(n_round_label = as.numeric(n_round_label)) 

# save the data frame as a csv file
write_csv(sessions_count, "../../data/processed/sessions_count.csv")

# remove data frame(s)
remove(sessions_count)
```

### bet_counts.csv

On each round, players can choose one of ten different bet sizes (from 0.25 euro to 20 euro). To characterize the betting patterns, for each player we count the number of rounds where they chose a certain bet size. The data are saved in the `bet_counts.csv` file, which contains the following variables.

* group: the group to which a player belongs, High-Risk or Low-Risk.
* id_player: the anonymous id of a player.
* stake: the amount of money a player bets on a round, from 0.25 till 20, in euro.
* n_round: the number of rounds where a player chose a certain bet size.
* n_round_total: the total number of rounds played by each player.
* prop: the relative proportion of choosing a certain bet size for each player, from 0 (%) till 100 (%).

```{r bet-patterns}
# get all the unique levels of bet sizes
bet_sizes <- unique(d$stake)
bet_sizes <- sort(bet_sizes, decreasing = F)

# for each player, count how often each bet size is selected
bets_count <- d %>%
  # turn the variable stake into a factor with all ten 10 possible levels
  mutate(stake = factor(stake, levels = bet_sizes)) %>%
  # for each player, count the number of rounds with a certain stake size
  count(group, id_player, stake, .drop = F, name = "n_round") %>%
  # calculate the total number of rounds for each player
  # and based on that, the relative proportion of choosing a certain stake size
  group_by(group, id_player) %>%
  mutate(
    n_round_total = sum(n_round), # the total number of rounds
    prop = n_round/n_round_total * 100  # the relative proportion of choosing a certain bet
  ) %>%
  ungroup()

# save the data frame as a csv file
write_csv(bets_count, "../../data/processed/bets_count.csv")

# remove data frame(s)
remove(bets_count)
```


## Stop

After each round, players can choose either to continue playing, or stop. Here we prepare data on whether the decision to stop is influenced by the prior outcome (win vs. loss) or not.

### stop.csv

For this analysis, we calculate the probability of winning on the last round of each session, as well as the overall probability of winning with all rounds combined, for each player. The idea is that if players made the decision to stop regardless of the outcome of a game, the overall probability of winning would be the same for the last round of each session and all rounds combined. However, if players were more likely to stop after a win, the last round would contain more wins than all rounds combined; conversely, if players were more likely to stop after a loss, the last round would contain more losses (i.e., a lower winning probability) than all rounds combined. The processed data are saved in the `stop.csv` file, with the following variables.

* group: the group to which a player belongs, High-Risk or Low-Risk.
* id_player: the anonymous id of a player.
* round_pos: whether a certain round is at the end of a session ('end') or all rounds combined ('overall').
* total_count: the total number of rounds in each condition, either at the end of a session or all rounds combined.
* win_count: the total number of rounds on which players won in each condition.
* win_prob: the probability of winning in each condition, calculated based on the previous two counts, from 0 (%) to 100 (%).

```{r stop}
# calculate the overall number of rounds and the probability of winning for each player
win_overall <- d %>%
  mutate(outcome = ifelse(win > 0, "win", "loss")) %>%
  group_by(group, id_player) %>%
  summarize(
    total_count = n(),
    win_count = sum(outcome == "win"),
    win_prob = win_count/total_count * 100
  ) %>%
  ungroup() %>%
  mutate(round_pos = "overall") %>%
  select(group, id_player, round_pos, total_count, win_count, win_prob)

# calculate the number of rounds that occur at the end of a session
# and the probability of winning in this subset of rounds
win_end_session <- d %>%
  mutate(outcome = ifelse(win > 0, "win", "loss")) %>%
  group_by(group, id_player, id_session) %>%
  mutate(
    # if a certain round has the largest id in a session
    # it means that round is the last one, and player ended the session afterwards
    round_pos = ifelse(id_round == max(id_round), "end", "in")
  ) %>%
  filter(round_pos == "end") %>%
  # for each player, calculate the number of rounds and wins at the end of each session
  group_by(group, id_player, round_pos) %>%
  summarize(
    total_count = n(),
    win_count = sum(outcome == "win"),
    win_prob = win_count/total_count * 100
  ) %>%
  select(group, id_player, round_pos, total_count, win_count, win_prob)

# combine the two data frames
stop <- win_overall %>%
  bind_rows(win_end_session) %>%
  arrange(group, id_player)

# save the data frame as a csv file
write_csv(stop, "../../data/processed/stop.csv")

# remove data frame(s)
remove(stop, win_overall, win_end_session)
```


## Stake change

Players can change their stakes between rounds. We explore whether the probability of changing stakes is influenced by prior outcome, and how players changed stakes when they did change stakes (again, as a function of prior outcome).

### stake_change.csv

Players can choose among 10 different levels of stake sizes (from 0.25 euro to 20 euro). Since we are interested in whether wins and losses may change the betting behavior, we exclude the first round of each session (i.e., these rounds have no prior outcomes). For the remaining data, we calculate the number of rounds that follow a win or a loss, and the probability of changing stakes.
We also calculate the average change in stakes (in euro; a positive change score indicates increasing stake, while a negative change score indicates decreasing stake). Since players can only select one from ten levels of stakes, as a second measure of stake change, we transform the stake amounts into stake levels, with 1 being the lowest level of stake (0.25 euro) and 10 being the highest level of stake (20 euro). A change score is then calculated based on the levels (rather than the original amounts) of stakes. We calculate the average change in stake sizes and stake levels both for all rounds, and only on rounds where players did change stakes.

The following variables are calculated and saved in the `stake_change.csv` file.

* group: the group to which a player belongs, High-Risk or Low-Risk.
* id_player: the anonymous id of a player.
* prev_outcome: the outcome on the previous round, win (win amount > 0) or loss (win amount = 0).
* n_round: the total number of rounds following a win or a loss.
* n_round_change: the number of rounds where a player changed the stake.
* stake_change_prop: the proportion of rounds where players changed the stake (%).
* stake_change_size_overall: the average change in stake sizes in all rounds. A positive value indicates an increase in stake (euro).
* stake_change_level_overall: the average change in stake levels in all rounds. A positive value indicates an increase in stake level.
* stake_change_size: the average change in stake sizes when players did change the stake.
* stake_change_level: the average change in stake levels when players did change the stake. 

```{r stake_change}
# get all the unique levels of bet sizes
bet_sizes <- unique(d$stake)
bet_sizes <- sort(bet_sizes, decreasing = F)

stake_change <- d %>%
  mutate(
    prev_outcome = ifelse(prev_win > 0, "win", "loss"),
    # create a new variable for stake level
    stake_level = factor(stake, levels = bet_sizes),
    stake_level = as.numeric(stake_level),
    # create a new variable for the previous stake
    prev_stake_level = lag(stake_level),
    # calculate the change in stakes between two rounds
    # a positive value indicates an increase in stake
    stake_change_size = stake - prev_stake,
    # do the same for the change in stake level
    stake_change_level = stake_level - prev_stake_level
  ) %>%
  filter(!is.na(prev_outcome)) %>%
  ungroup()

# calculate the probability of changing stakes and the average change
# in stake sizes and levels with all rounds
stake_change_prop <- stake_change %>%
  group_by(group, id_player, prev_outcome) %>%
  summarize(
    # the total number of rounds across all sessions
    n_round = n(),
    # the total number of rounds when stakes did change
    n_round_change = sum(stake_change_size != 0),
    # probability of changing stakes
    stake_change_prop = mean(stake_change_size != 0) * 100,
    # the overall change in stake sizes and levels
    stake_change_size_overall = mean(stake_change_size),
    stake_change_level_overall = mean(stake_change_level)
  ) %>%
  ungroup()

# calculate the changes in stakes when players did change stakes
stake_change_sizes <- stake_change %>%
  # only include rounds where players did change stakes
  filter(stake_change_size != 0) %>%
  group_by(group, id_player, prev_outcome) %>%
  summarize(
    stake_change_size = mean(stake_change_size),
    stake_change_level = mean(stake_change_level),
  ) %>%
  ungroup()

# combine both data frames
stake_change_combined <- stake_change_prop %>%
  full_join(stake_change_sizes, by = c("group", "id_player", "prev_outcome"))

# save the data frame as a csv file
write_csv(stake_change_combined, "../../data/processed/stake_change.csv")

remove(stake_change, stake_change_prop, stake_change_sizes, stake_change_combined)
```

## Speed of play

Next we examine the speed of play, once players have decided to continue playing. For this analysis, we focus on the RT of putting in the first column as an indicator of the speed of play. For this analysis, we further exclude rounds where the RT of putting in the first column was at least 5000 milliseconds. RTs below 0 (potentially a recording error) are also excluded. 

### rt.csv

```{r rt-data-exclusion}
# number of rounds before exclusion
before <- nrow(d)

d_after <- d %>%
  # the rt1 in the data registers the duration between the end of the previous game
  # and the moment when players put in the first column in the current game.
  # by calculating the difference between 'rt1' and 'rt_start',
  # we get the rt of the putting in the first column,
  # which is between starting the current game and putting in the first column.
  mutate(rt = rt1 - rt_start) %>%
  filter(
    # exclude the first round of each session
    id_round !=1, 
    # only keep trials where the RT of putting in the first column
    # is between 0 and 5000 milliseconds
    rt < 5000, rt >= 0 
  ) 

# calculate number of rounds after the exclusion
after <- nrow(d_after)

# calculate the proportion of data exclusion
removed <- (before - after)/before * 100
```

Players in total played `r before` rounds. `r before - after` rounds were excluded after applying the exclusion criteria as listed above (`r round(removed, 1)`\% of all rounds). For the remaining rounds, we calculate the number of rounds, and the mean RTs following wins and losses respectively. Since players differed in how quickly they generally responded, to control for the potential effects of general responding speeds, we further calculate the z score of each RT within each player. The mean RT z scores are further calculated for each player. The processed data are saved in the `rt.csv` file. The following variables are included.

* group: the group to which a player belongs, High-Risk or Low-Risk.
* id_player: the anonymous id of a player.
* prev_outcome: the outcome on the previous trial, win (win amount > 0) or loss (win amount = 0).
* n_round: the number of remaining rounds following a win or a loss.
* rt_mean: the mean response time (of putting in the first column) after a win or a loss, in milliseconds.
* rt_z_mean: the mean of response time z score (of putting in the first column) after a win or a loss.

```{r rt-data}
rt_data <- d_after %>%
  # calculate the following variables for each player
  group_by(group, id_player) %>%
  mutate(
    rt_z = scale(rt),
    prev_outcome = ifelse(prev_win > 0, "win", "loss") # the outcome of the previous round
  ) %>%
  # for rounds following wins and losses, calculate the total number
  # and the mean RTs (raw scores as well as z scores)
  group_by(group, id_player, prev_outcome) %>%
  summarize(
    n_round = n(), # number of rounds
    rt_mean = mean(rt), # the mean rt 
    rt_z_mean = mean(rt_z) # the mean rt z score
  )

# save the data frame as a csv file
write_csv(rt_data, "../../data/processed/rt.csv")

remove(d_after, rt_data, d)
```
